---
title: "Coding and Results"
author: "McKenzie"
date: "2023-07-18"
output:
  word_document: default
  html_document: default
---

##CLEANING 

*Loading Packages*
```{r,include=FALSE,echo=FALSE}
##libraries
library(SmartEDA)
library(tidygeocoder)
library(dplyr)
library(xts) #Load package
library(sf)
library(ggplot2)
library(tidyverse)
library(forecast)
library(rvest)
library(trend)
library(lubridate)
library(tseries)
library(TTR)
```


* The shelter data is being pulled from the city of Toronto's [Daily Shelter & Overnight Service Occuapncy & Capacity](https://open.toronto.ca/dataset/daily-shelter-overnight-service-occupancy-capacity/) dataset on Toronto's open portal website. 

* The dataset is split into 3 files,1 for each year. For the 2023 year file, the data source is updated daily. 

```{r,echo=FALSE}
##Loading data and cleaning date field for union
 
shelter2021<-read.csv('C:\\Users\\User\\Desktop\\Big Data Analytics Program\\Project\\Tentative Project Datasets\\Shelter\\Daily Shelter Occupancy\\Current\\Daily shelter overnight occupancy (2021).csv')
shelter2021$OCCUPANCY_DATE<-as.Date(shelter2021$OCCUPANCY_DATE, format = "%y-%m-%d")

shelter2022<-read.csv('C:\\Users\\User\\Desktop\\Big Data Analytics Program\\Project\\Tentative Project Datasets\\Shelter\\Daily Shelter Occupancy\\Current\\Daily shelter overnight occupancy (2022).csv')
shelter2022$OCCUPANCY_DATE<-as.Date(shelter2022$OCCUPANCY_DATE, format = "%y-%m-%d")

shelter2023<-read.csv('C:\\Users\\User\\Downloads\\Daily shelter overnight occupancy (2).csv')
shelter2023$OCCUPANCY_DATE<-as.Date(shelter2023$OCCUPANCY_DATE, format = "%Y-%m-%d")

#Union for complete dataset
shltr<-rbind(shelter2021,shelter2022,shelter2023)
rm(shelter2021,shelter2022,shelter2023)

##Creating OVER_OCCUPIED field. Identifying programs that are over occupied exclusive of the capacity type
##NEED TO UPDATE TO TAKE INTO CONSIDERATION SHELTERS OPERATING AT OVER CAPACITY~~~~~~~
shltr$OVER_OCCUPIED<-ifelse(shltr$OCCUPANCY_RATE_ROOMS==100|shltr$OCCUPANCY_RATE_BEDS==100,1,0)
shltr$OVER_OCCUPIED<-ifelse(is.na(shltr$OVER_OCCUPIED)==TRUE,0,1)

##Creating ADDRESS field. It is a program's full address. 
shltr$ADDRESS<-paste(shltr$LOCATION_ADDRESS,shltr$LOCATION_CITY,shltr$LOCATION_PROVINCE)

##OCCUPANCY_RATES. Creating one field for occupancy rates inclusive of capacity type
shltr$OCCUPANCY_RATE<-ifelse(is.na(shltr$OCCUPANCY_RATE_BEDS),shltr$OCCUPANCY_RATE_ROOMS,shltr$OCCUPANCY_RATE_BEDS)

##Drop redundant fields
shltr<-shltr[,c(1:20,33,34,35)]


```


*Dealing with Missing Values*

```{r,echo=FALSE}
ExpData(shltr,type=2)
```

* The proportion of missing values is very small. Most of the missing values are for fields related to location. Since the fields that contain missing values are non-numerical I am going to convert all missing values to "unknown". 

```{r,echo=FALSE}
#any missing values for measures? 
anyNA(shltr$OCCUPANCY_RATE)
anyNA(shltr$OVER_OCCUPIED)
```

```{r,echo=FALSE}
#The only instances of missing values appear as blanks in location related fields. I am turning all blanks to NAs 
shltr[shltr==""]<-"Unknown"
shltr[is.na(shltr)]<-"Unknown"
```


```{r,echo=FALSE}
#Need to trim leading white spaces from city field
shltr$LOCATION_CITY<-trimws(shltr$LOCATION_CITY)

```

*Additional geographic information*

* Part of my analysis will involve understanding shelter related metrics across different neighborhood. 

* The following code brings in location related characteristics. The [census tract boundary file](https://www12.statcan.gc.ca/census-recensement/alternative_alternatif.cfm?l=eng&dispext=zip&teng=lct_000b21a_e.zip&k=%20%20%20%2013089&loc=//www12.statcan.gc.ca/census-recensement/2021/geo/sip-pis/boundary-limites/files-fichiers/lct_000b21a_e.zip) comes from statistics Canada and the city of Toronto's [Neighbourhood file.](https://open.toronto.ca/dataset/neighbourhoods/)

```{r,echo=FALSE}

##Creating a list of addresses for each shelter program. These addresses will be used to bring Census tract/neighbourhoods for each program. 
Addresses<-as.data.frame(unique(paste(shltr$LOCATION_ADDRESS,shltr$LOCATION_CITY,shltr$LOCATION_PROVINCE)))
#Cleaning Column Name
colnames(Addresses)<-'Addresses'

#Using the tidygeocoder package to bring in lat and longs to each address for neighbourhood identification
lat_longs<-Addresses %>% 
  geocode(Addresses)
#Removing postal codes that did not return a lat/long
lat_longs<-lat_longs %>%
  filter(is.na(lat)==FALSE & lat_longs$Addresses!='')
#Create a point geometric field using the st package
lat_longs<-lat_longs %>% st_as_sf(coords=c('long','lat'))
lat_longs<-st_set_crs(lat_longs,4326)

#Raading the census tract boundary file from statistics canada. 
# Provide the link 

CT<-read_sf("C:\\Users\\User\\Desktop\\CTest\\lct_000a21a_e.shp")
#Filter for Ontario Province
CT<-CT %>% filter(PRUID=='35')
CT<-st_transform(CT,crs=4326)

#Toronto Neighbourhoods Profile. To identify a Toronto Neighbourhood map geometric point to the boundary file found on toronto open data portal
TNei<-read_sf('C:\\Users\\User\\Desktop\\Big Data Analytics Program\\Project\\Tentative Project Datasets\\Neighbourhoods\\Boundary File\\Neighbourhoods - 4326.shp')
#Table identifying each CT Id for each shelter
TorCTs<-st_join(lat_longs,CT)

#Decide which fields to keep, bring neighbourhood information to address data source
TorCTs<-st_join(TorCTs,TNei)
TorCTsMap<-TorCTs
TorCTs<-TorCTs[,c(1,15,16)]

##Bring Neighbourhood information to the shltr data set. 
shltr<-left_join(shltr,TorCTs,by=c('ADDRESS' ='Addresses'))
colnames(shltr)[c(24,25)]<-c('Neighbourhood','Improv_Status_Area')


##Adding in weekday
shltr$WeekDay<-weekdays(shltr$OCCUPANCY_DATE)
shltr$Month<-format(shltr$OCCUPANCY_DATE,'%b')
shltr$ShltrSize<-round(shltr$SERVICE_USER_COUNT/(shltr$OCCUPANCY_RATE/100),digits = 0)
```


**Creating Measures of Interest** 

```{r,echo=FALSE}
#Defining the index for time for the time series object
minDate<-min(shltr$OCCUPANCY_DATE)
maxDate<-max(shltr$OCCUPANCY_DATE)
indy<-seq(minDate,maxDate,by='day')
```

*Creating measures* 

* Note all of these measures were used in the analysis. 
```{r, echo=FALSE}
##Daily Intakes Aggregated for the Toronto Region. 
NTakeShltr<-aggregate(SERVICE_USER_COUNT ~ OCCUPANCY_DATE ,data=shltr,FUN=sum)
colnames(NTakeShltr)[2]<-'TotalUsers'
##Creating xts(time series) object
NTakeShltrVec<-NTakeShltr$TotalUsers
xtsNTakeShltr<-xts(NTakeShltrVec,order.by = indy)
colnames(xtsNTakeShltr)<-'Total_Intakes'
rm(NTakeShltrVec)

##Daily Count of programs operating
OpenShltr<-aggregate(PROGRAM_ID ~ OCCUPANCY_DATE ,data=shltr,FUN=length)
colnames(OpenShltr)[2]<-'TotalOpenPrograms'
#xts object
OpenShltrVec<-OpenShltr$TotalOpenPrograms
xtsOpenShltr<-xts(OpenShltrVec,order.by=indy)
colnames(xtsOpenShltr)<-'TotalOpenPrograms'
rm(OpenShltrVec)

##Over Capacity Programs
OverCapShltr<-aggregate(OVER_OCCUPIED ~ OCCUPANCY_DATE ,data=shltr,FUN=sum)
colnames(OverCapShltr)[2]<-'TotalOverCapacityShltrs'
#xts object
OverCapShltrVec<-OverCapShltr$TotalOverCapacityShltrs
xtsOverCapShltr<-xts(OverCapShltrVec,order.by = indy)
colnames(xtsOverCapShltr)<-'TotalOverCapacityShltrs'
rm(OverCapShltrVec)

#Proportion of shltrs operating at over capacity
OverCapPropShltr<-left_join(OpenShltr,OverCapShltr,by = 'OCCUPANCY_DATE')
OverCapPropShltr$OverCapRate<-round((OverCapPropShltr$TotalOverCapacityShltrs/OverCapPropShltr$TotalOpenPrograms)*100,digits = 2)
OverCapPropShltr<-OverCapPropShltr[,c(1,4)]
#xts object
OverCapPropShltrVec<-OverCapPropShltr$OverCapRate
xtsOverCapPropShltr<-xts(OverCapPropShltrVec,order.by = indy)
colnames(xtsOverCapPropShltr)<-'OverCapRate'

##Average Capacity Rate
AvgOccPerShltr<-aggregate(OCCUPANCY_RATE ~ OCCUPANCY_DATE,data=shltr,FUN=mean)
AvgOccPerShltr$OCCUPANCY_RATE<-round(AvgOccPerShltr$OCCUPANCY_RATE,digits = 2)
AvgOccPerShltr$OccupancyRate<-NULL
#xts object
AvgOccPerShltrVec<-AvgOccPerShltr$OCCUPANCY_RATE
xtsAvgOccPerShltr<-xts(AvgOccPerShltrVec,order.by = indy)
colnames(xtsAvgOccPerShltr)<-'OCCUPANCY_RATE'

##Average Size of shelter programs
AvgShltrSze<-aggregate(shltr$ShltrSize ~ OCCUPANCY_DATE ,data=shltr,FUN=mean)
```

##Neighbourhood Improvement Area

```{r,echo=FALSE}
#Clean Neighbourhood improvement area column
unique(shltr$Improv_Status_Area)
shltr$Improv_Status_Area<-ifelse(shltr$Improv_Status_Area!='Neighbourhood Improvement Area' | is.na(shltr$Improv_Status_Area)==TRUE ,'Not Improvment Area','Neighbourhood Improvement Area')

###Aggregate on improvement area and occupancy date to get the daily mean occupancy rate, total programs operating and total programs that hit capacity
ImprAvgCap<- shltr %>%
  group_by(OCCUPANCY_DATE, Improv_Status_Area) %>% 
  summarise(AVG_CAPACITY = mean(OCCUPANCY_RATE),
            TotalProgrms = n(),
            Hit_Capacity = sum(OVER_OCCUPIED))
##Getting the proportion of shelters at or over capacity.
ImprAvgCap$OverCapacityRate <- round((ImprAvgCap$Hit_Capacity / ImprAvgCap$TotalProgrms)*100,digits = 2)
```

```{r}

##Creating line graph for average daily capacity rates by neighbourhood improvement area
ggplot(ImprAvgCap,aes(x=as.Date(OCCUPANCY_DATE), y=AVG_CAPACITY,color=Improv_Status_Area ))+
  geom_line()+
  labs(x='Date', y='Daily Average Capacity Rate', color='Improvement_area_status')+
  facet_wrap(~Improv_Status_Area,ncol = 1);

##Creating a line graph for the proportion of shelters operating at or over capacity
ggplot(ImprAvgCap,aes(x=as.Date(OCCUPANCY_DATE), y=OverCapacityRate,color=Improv_Status_Area ))+
  geom_line()+
  labs(x='Date', y='Daily Average OVer Capacity Rate', color='Improvement_area_status')+
  facet_wrap(~Improv_Status_Area,ncol = 1)

```

```{r}
##Aggregating to get the average of the average capacity rates for neighbourhood improvement areas and the average proporiton of shelters operating at or over capacity
ImprOverallAvgCap<-ImprAvgCap %>%
  group_by(Improv_Status_Area) %>%
  summarise(AVG_CAPACITY = round(mean(AVG_CAPACITY),digits=2),
            AVG_OVER_CAPACITY_RATE = round(mean(OverCapacityRate),digits = 2))

##Creating a barchart to compare averages
ggplot(ImprOverallAvgCap,aes(x=Improv_Status_Area, y=AVG_CAPACITY,fill=Improv_Status_Area ))+
  geom_bar(stat='identity')+
  theme(axis.text = element_text(angle=90, hjust=1))+
  geom_text(aes(label = AVG_CAPACITY),vjust=1.5,color='black')+
  labs(title = 'Average Capacity Rate',x='Improvement Area Status',color = 'Improvement Area Status');
  
ggplot(ImprOverallAvgCap,aes(x=Improv_Status_Area, y=AVG_OVER_CAPACITY_RATE,fill=Improv_Status_Area ))+
  geom_bar(stat='identity')+
  theme(axis.text = element_text(angle=90, hjust=1))+
  geom_text(aes(label = AVG_OVER_CAPACITY_RATE),vjust=1.5,color='black')+
  labs(title = 'Average Over Capacity Rate',x='Improvement Area Status',color = 'Improvement Area Status')
```


```{r}
##Kruskal Wallis test 
###Test for comparing average capacity rates
kruskal.test(AVG_CAPACITY ~ Improv_Status_Area,data=ImprAvgCap)
##Test for comparing the average of proportion of shelters operating at or over capacity. 
kruskal.test(OverCapacityRate ~ Improv_Status_Area ,data=ImprAvgCap)
```

## Average Capacity Rate Trend**

```{r,echo=FALSE}
# Convert the xts object to a data frame
AvgOccPerShltrPlt <- data.frame(Date = index(xtsAvgOccPerShltr), Value = coredata(xtsAvgOccPerShltr))
#Creating label to identify dates where we see the maximum shelter intakes
max_date <- index(xtsAvgOccPerShltr)[which.max(coredata(xtsAvgOccPerShltr))]
#Creating label to identify dates where we see the minimum shelter intakes
min_date <- index(xtsAvgOccPerShltr)[which.min(coredata(xtsAvgOccPerShltr))]

# Create the plot
ggplot(data = AvgOccPerShltrPlt, aes(x = Date, y = OCCUPANCY_RATE)) +
  geom_line()+
#Adding max and min label dates  
  geom_text(data = subset(AvgOccPerShltrPlt, Date %in% c(max_date, min_date)),
            aes(label = as.character(Date), vjust = ifelse(Date == max_date, -0.5, 0.5)),
            show.legend = FALSE)+
#Adding points to id min and max dates  
  geom_point(data = subset(AvgOccPerShltrPlt, Date %in% c(max_date, min_date)),
             aes(color = ifelse(Date == max_date, "blue",'red')),
             size = 3)+
  labs(title = "Average Daily Shelter Capacity Rates") +
  xlab("Date") +
  ylab("Value")+
  theme(legend.position = 'none');plot.ts(SMA(xtsAvgOccPerShltr$OCCUPANCY_RATE,n=5))

```



|Metric|Daily Capacity Rate|
|--------|------------|
|Mean|`r round(mean(xtsAvgOccPerShltr),digits=2)`|
|Median |`r round(median(xtsAvgOccPerShltr),digits=2)` |
|Standard Deviation|`r round(sd(xtsAvgOccPerShltr),digits=2)`|
|Minimum|`r round(min(xtsAvgOccPerShltr),digits=2)`|
|Maximum|`r round(max(xtsAvgOccPerShltr),digits=2)`|
|Range |`r round(min(xtsAvgOccPerShltr),digits=2)` - `r round(max(xtsAvgOccPerShltr),digits=2)`|

**Capacity Rate Distribution**

```{r,echo=FALSE}
ggplot(xtsAvgOccPerShltr,
       aes(sample=OCCUPANCY_RATE))+
  geom_qq()+
  geom_qq_line()

```
**ACF PLOT for Average Daily Capacity Rates**

```{r,echo=FALSE}
acf(ts(xtsAvgOccPerShltr))
```

**Differenced plot**

```{r,echo=FALSE}
plot.ts(diff(xtsAvgOccPerShltr))
```

**Average Daily Capacity Rate: Stationairy** 

```{r,echo=FALSE}
kpss.test(ts(xtsAvgOccPerShltr))##KPSS TEST
adf.test(ts(xtsAvgOccPerShltr))##ADF TEST
```

**Average Capacity RAte: Trend Assessment**

```{r,echo=FALSE}
mkAvgOccShlter<-mk.test(ts(xtsAvgOccPerShltr$OCCUPANCY_RATE))
print(mkAvgOccShlter)
snsAvgOccShltr<-sens.slope(ts(xtsAvgOccPerShltr$OCCUPANCY_RATE))
print(snsAvgOccShltr)
```


### Capacity Rate Forcasting

```{r,echo=FALSE}
##Creating test and training sets.
#For forecasting I will convert my xts object to a ts object. 
tsAvgOccPerShltr<-ts(xtsAvgOccPerShltr)
##Training set is all the dates in the time series except the last 30 days (a month)
trnAvgOccPerShltr<-window(tsAvgOccPerShltr,end = end(index(tsAvgOccPerShltr)[1:(nrow(tsAvgOccPerShltr)-31)]))
##Test
tstAvgOccPerShltr<-window(tsAvgOccPerShltr,start=c(end(tsAvgOccPerShltr)[1] - 30, end(tsAvgOccPerShltr)[2]))

```

```{r,echo=FALSE}
#Naive Test
fcNaiAvgOcc<-naive(trnAvgOccPerShltr,h=30)
autoplot(fcNaiAvgOcc)+
  autolayer(tstAvgOccPerShltr,series='Test Date')
```

```{r,echo=FALSE}
##Random walk with drift
fcRwdAvgOcc<-rwf(trnAvgOccPerShltr,h=30,drift=TRUE)
autoplot(fcRwdAvgOcc)+
  autolayer(tstAvgOccPerShltr,series='Random Walk With Drift')
```

```{r,echo=FALSE}
#Simple Exponential Smoothing 
fcSesAvgOcc<-ses(trnAvgOccPerShltr,h=30)
autoplot(fcSesAvgOcc)+
  autolayer(tstAvgOccPerShltr,series = "SES Test")
forecast(fcSesAvgOcc,h=30)
```

```{r,echo=FALSE}
##ARIMA
ArAvgOcc<-auto.arima(trnAvgOccPerShltr)
fcArAvgOcc<-forecast(ArAvgOcc,h=31)
autoplot(fcArAvgOcc)+
  autolayer(tstAvgOccPerShltr,series='ARIMA Test')
```


##Evaluating Model Performance: Capacity RAte

```{r,echo=FALSE}
##Pulling MAPE Valeus and assigning to vectors
MapeARMAvgOCC<-accuracy(fcArAvgOcc,tstAvgOccPerShltr)[2,5]
MapeSesAvgOCC<-accuracy(fcSesAvgOcc,tstAvgOccPerShltr)[2,5]
MapeNaiAvgOcc<-accuracy(fcNaiAvgOcc,tstAvgOccPerShltr)[2,5]
MapeRWDAvgOCC<-accuracy(fcRwdAvgOcc,tstAvgOccPerShltr)[2,5]

MapeValues<-c(round(MapeARMAvgOCC,digits = 3),round(MapeSesAvgOCC,digits = 3) ,round(MapeNaiAvgOcc,digits = 3),round(MapeRWDAvgOCC,digits = 3))
#Barplot to compare 
barplot(MapeValues,names.arg = c('ARIMA','SES','Naive','Random Walk With Drift'))
text(x=1:length(MapeValues),y=MapeValues,labels = MapeValues,pos=1)

```

**Checking Risduals**
```{r,echo=FALSE}
checkresiduals(fcArAvgOcc)
```

###Trend Segmentation: Average Capacity Rate


```{r,echo=FALSE}
##Creating time series set just for after the peak
AftrPeak<-xtsAvgOccPerShltr[index(xtsAvgOccPerShltr)>index(xtsAvgOccPerShltr)[which.max(coredata(xtsAvgOccPerShltr))],]
plot.ts(AftrPeak)
```
**Distribution of Capacity Rate After PEak** 
```{r}
qplot(AftrPeak)
```

**After Peak: Stationary Test**

```{r,echo=FALSE}
mk.test(ts(AftrPeak))
```


```{r,echo=FALSE}
kpss.test(ts(AftrPeak));adf.test(ts(AftrPeak))
```


###Proportion of shelters operating at or over capacity

```{r,echo=FALSE}
# Convert the xts object to a data frame
OverCapPropShltrPlt <- data.frame(Date = index(xtsOverCapPropShltr), Value = coredata(xtsOverCapPropShltr))
#Creating label to identify dates where we see the maximum shelter intakes
max_date <- index(xtsOverCapPropShltr)[which.max(coredata(xtsOverCapPropShltr))]
#Creating label to identify dates where we see the minimum shelter intakes
min_date <- index(xtsOverCapPropShltr)[which.min(coredata(xtsOverCapPropShltr))]

# Create the plot
ggplot(data = OverCapPropShltrPlt, aes(x = Date, y = OverCapRate)) +
  geom_line()+
#Adding max and min label dates  
  geom_text(data = subset(OverCapPropShltrPlt , Date %in% c(max_date, min_date)),
            aes(label = as.character(Date), vjust = ifelse(Date == max_date, -0.5, 0.5)),
            show.legend = FALSE)+
#Adding points to id min and max dates  
  geom_point(data = subset(OverCapPropShltrPlt , Date %in% c(max_date, min_date)),
             aes(color = ifelse(Date == max_date, "blue",'red')),
             size = 3)+
  labs(title = "% of Shelters at or over Capacity") +
  xlab("Date") +
  ylab("Value")+
  theme(legend.position = 'none');plot.ts(SMA(xtsOverCapPropShltr,n=5))
```


**Distribution: At or Over Capacity**

```{r,echo=FALSE}
ggplot(xtsOverCapPropShltr,
       aes(sample=OverCapRate))+
  geom_qq()+
  geom_qq_line()
```


**ACF PLOT: At or Over Capacity Rate**

```{r,echo=FALSE}
acf(ts(xtsOverCapPropShltr))
```

**Stationary Assessment: At or Over Capacity** 

```{r,echo=FALSE}
kpss.test(ts(xtsOverCapPropShltr))##KPSS TEST
adf.test(ts(xtsOverCapPropShltr)) ##ADF TEST
```

**Trend Assessment: At or Over Capacity**

```{r,echo=FALSE}
mkOverCapPropShltr<-mk.test(ts(xtsOverCapPropShltr))
print(mkOverCapPropShltr)
snsOverCapPropShltr<-sens.slope(ts(xtsOverCapPropShltr$OverCapRate))
print(snsOverCapPropShltr)
```

**Forecasting: At or Over Capacity**

```{r,echo=FALSE}
##Creating test and training sets.
#For forecasting I will convert my xts object to a ts object. 
tsOverCapPropShltr<-ts(xtsOverCapPropShltr)
##Training is all the dates in the time series except the last 30 days (a month)
trnOverCapPropShltr<-window(tsOverCapPropShltr,end = end(index(tsOverCapPropShltr)[1:(nrow(tsOverCapPropShltr)-31)]))
##Test
tstOverCapPropShltr<-window(tsOverCapPropShltr,start=c(end(tsOverCapPropShltr)[1] - 30, end(tsOverCapPropShltr)[2]))
```

```{r,echo=FALSE}
#Naive Test
fcNaiOverCap<-naive(trnOverCapPropShltr,h=30)
autoplot(fcNaiOverCap)+
  autolayer(tstOverCapPropShltr,series='Test Date')
```

```{r,echo=FALSE}
##Random walk with drift
fcRwdOverCap<-rwf(trnOverCapPropShltr,h=30,drift=TRUE)
autoplot(fcRwdOverCap)+
  autolayer(tstOverCapPropShltr,series='Random Walk With Drift')
```

```{r,echo=FALSE}
#Simple Exponential Smoothing 
fcSesOverCap<-ses(trnOverCapPropShltr,h=30)
autoplot(fcSesOverCap)+
  autolayer(tstOverCapPropShltr,series = "SES Test")

```

```{r,echo=FALSE}
#ARIMA
ArOverCap<-auto.arima(trnOverCapPropShltr)
fcArOverCap<-forecast(ArOverCap,h=31)
autoplot(fcArOverCap)+
  autolayer(tstOverCapPropShltr,series='ARIMA Test')

```

**Evaluating Model Performance: At or Over Capacity**

```{r,echo=FALSE}
##Pulling MAPE values
MapeARMOverCap<-accuracy(fcArOverCap,tstOverCapPropShltr)[2,5]
MapeSesOverCap<-accuracy(fcSesOverCap,tstOverCapPropShltr)[2,5]
MapeNaiOverCap<-accuracy(fcNaiOverCap,tstOverCapPropShltr)[2,5]
MapeRWDOverCap<-accuracy(fcRwdOverCap,tstOverCapPropShltr)[2,5]

MapeValues<-c(round(MapeARMOverCap,digits = 3),round(MapeSesOverCap,digits = 3) ,round(MapeNaiOverCap ,digits = 3),round(MapeRWDOverCap ,digits = 3))

barplot(MapeValues,names.arg = c('ARIMA','SES','Naive','Random Walk With Drift'))
text(x=1:length(MapeValues),y=MapeValues,labels = MapeValues,pos=1)

```

**Residuals: At or Over Capacity**

```{r,echo=FALSE}
checkresiduals(fcArOverCap)
```


**Correlation: Capacity Rate ~ Over Capacity**

```{r,echo=FALSE}
plot(as.vector(xtsAvgOccPerShltr$OCCUPANCY_RATE), as.vector(xtsOverCapPropShltr$OverCapRate))
```
```{r,echo=FALSE}
cor(as.vector(xtsAvgOccPerShltr$OCCUPANCY_RATE), as.vector(xtsOverCapPropShltr$OverCapRate),method = 'pearson' )
```


##Refugee Programs: Capacity Rate

**Looking at those more susceptible to homelessness** 

# Creating Refugee identification column
```{r,echo=FALSE}
shltr$RfugPrgm<-ifelse(grepl('Refugee',shltr$PROGRAM_NAME)|shltr$PROGRAM_AREA=='Temporary Refugee Response' ,'Refugee','Other')
```

**Trend: Refugee Programs vs non-refugee programs**

```{r,echo=FALSE}
RefuAvgOcc<-shltr %>%
  group_by(OCCUPANCY_DATE,RfugPrgm) %>%
  summarise(AVG_CAPACITY = mean(OCCUPANCY_RATE),
            Programs = n(),
            AverageNTakes = mean(SERVICE_USER_COUNT),
            WeightedAvgUsers = weighted.mean(SERVICE_USER_COUNT,ShltrSize),
            WeightAvgOcc =weighted.mean(OCCUPANCY_RATE,ShltrSize),
            TotalUsers = sum(SERVICE_USER_COUNT))

ggplot(RefuAvgOcc,aes(x=as.Date(OCCUPANCY_DATE),y=AVG_CAPACITY))+
  geom_line()+
  facet_wrap(~RfugPrgm)

```

**Testing Averages**

```{r,echo=FALSE}
AvgOccRefug<- RefuAvgOcc %>% filter(RfugPrgm=='Refugee')
xtsAvgOccRefug<-xts(x=AvgOccRefug$AVG_CAPACITY,order.by = AvgOccRefug$OCCUPANCY_DATE)
AvgOccOthr<- RefuAvgOcc %>% filter(RfugPrgm!='Refugee')
xtsAvgOccOthr<-xts(x=AvgOccOthr$AVG_CAPACITY,order.by = AvgOccOthr$OCCUPANCY_DATE)
```

```{r,echo=FALSE}
wilcox.test(as.vector(AvgOccRefug$AVG_CAPACITY),as.vector(AvgOccOthr$AVG_CAPACITY))
```


**Correlation between refugee programs and other programs**
```{r,echo=FALSE}
plot(as.vector(xtsAvgOccRefug), as.vector(xtsAvgOccOthr))
cor(as.vector(xtsAvgOccRefug), as.vector(xtsAvgOccOthr),method = 'pearson' )
```

**Differenced plots in refugee capacity rates and other programs**
```{r,echo=FALSE}
plot.ts(diff(xtsAvgOccOthr));plot.ts(diff(xtsAvgOccRefug))

```

**ACF PLOTS: Refugee and non-refugee programs**

```{r,echo=FALSE}
acf(ts(xtsAvgOccRefug));acf(ts(xtsAvgOccOthr))
```

**Stationary assessment: Refugee programs**

```{r,echo=FALSE}
kpss.test(ts(xtsAvgOccRefug))
adf.test(ts(xtsAvgOccRefug))
```

**Trend Assessment: Refugee Programs**

```{r,echo=FALSE}
mkAvgOccRefug<-mk.test(ts(xtsAvgOccRefug))
print(mkAvgOccRefug)
snsAvgOccRefug<-sens.slope(ts(xtsAvgOccRefug))
print(snsAvgOccRefug)

```


**Forecasting: Refugee Programs** 

```{r,echo=FALSE}
##Creating test and training sets.
#For forecasting I will convert my xts object to a ts object. 
tsAvgOccRefug<-ts(xtsAvgOccRefug)
##Training is all the dates in the time series except the last 30 days (a month)
trnAvgOccRefug<-window(tsAvgOccRefug,end = end(index(tsAvgOccRefug)[1:(nrow(tsAvgOccRefug)-31)]))
##Test
tstAvgOccRefug<-window(tsAvgOccRefug,start=c(end(tsAvgOccRefug)[1] - 30, end(tsAvgOccRefug)[2]))
```

```{r}
#Naive Test
fcNaiAvgOccRefug<-naive(trnAvgOccRefug,h=30)
autoplot(fcNaiAvgOccRefug)+
  autolayer(tstAvgOccRefug,series='Test Date')
```

```{r,echo=FALSE}
##Random walk with drift
fcRwdAvgOccRefug<-rwf(trnAvgOccRefug,h=30,drift=TRUE)
autoplot(fcRwdAvgOccRefug)+
  autolayer(tstAvgOccRefug,series='Random Walk With Drift')
```

```{r,echo=FALSE}
#Simple Exponential Smoothing 
fcSesAvgOccRefug<-ses(trnAvgOccRefug,h=30)
autoplot(fcSesAvgOccRefug)+
  autolayer(tstAvgOccRefug,series = "SES Test")

```


```{r,echo=FALSE}
##ARIMA
ArAvgOccRefug<-auto.arima(trnAvgOccRefug)
fcArAvgOccRefug<-forecast(ArAvgOccRefug,h=31)
autoplot(fcArAvgOccRefug)+
  autolayer(tstAvgOccRefug,series='ARIMA Test')

```

**Evaluating Model Performance: Refugee Capacity Rates**

```{r,echo=FALSE}
MapeARMOverCap<-accuracy(fcArAvgOccRefug,tstAvgOccRefug)[2,5]
MapeSesOverCap<-accuracy(fcSesAvgOccRefug,tstAvgOccRefug)[2,5]
MapeNaiOverCap<-accuracy(fcNaiAvgOccRefug, tstAvgOccRefug)[2,5]
MapeRWDOverCap<-accuracy(fcRwdAvgOccRefug,tstAvgOccRefug)[2,5]

MapeValues<-c(round(MapeARMOverCap,digits = 3),round(MapeSesOverCap,digits = 3) ,round(MapeNaiOverCap ,digits = 3),round(MapeRWDOverCap ,digits = 3))

barplot(MapeValues,names.arg = c('ARIMA','SES','Naive','Random Walk With Drift'))
text(x=1:length(MapeValues),y=MapeValues,labels = MapeValues,pos=1)
```

```{r,echo=FALSE}
checkresiduals(fcRwdAvgOccRefug)

```





